---
title: "679 Final Project"
author: "Team 9"
date: "4/12/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(plyr)
library(reshape2)
library(tidyr)
library(viridis)
library(psych)
library(rstanarm)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
DF <- read.csv("transformed.csv", encoding="UTF-8")
describe(DF)
```

```{r}
colnames(DF) <- c("Study.ID","Sex",                                    "Year.of.Diagnosis","Age.at.Diagnosis","Race",                                       "Insurance","SEER.Registry","X...9th.Grade.Education",                         "X...High.School.Education","X...Bachelors.Education",                         "X..Persons.Below.Poverty","X..Unemployed.ACS.2013.2017",                     "Median.Household.Income",                      "X..Language.isolation.ACS.2013.2017..households.","Site",                              "Subsite","AJCC.7.Stage","Size",                                            "Lymph.Nodes","Mets","Cause.of.Death",                                  "Surgery.Performed.","Surgery.Decision","Radiation",                                    "Chemotherapy")
colnames(DF)
```

# Exploratory Data Analysis

## Cancer Selection

  There are seven kinds of cancer with different stages in this dataset. To simplify our research, we decide to choose only one kind of cancer to do the analysis. 
  So, we firstly explore the top 10 cancers(including the stages) that the patients are most likely to give up the surgery. Giving up the surgery can suggest more that the patients are affected by some factors such as biases in the society.

```{r message=FALSE}
# Surgery.Decision
df_cancer_stage <- DF %>% 
  dplyr::select(Site,AJCC.7.Stage,Surgery.Decision,Study.ID) %>% 
  unite("Cancer_Stage",Site,AJCC.7.Stage,sep = "-") %>% 
  dplyr::group_by(Cancer_Stage,Surgery.Decision) %>% 
  dplyr::summarise(num = n()) %>% ungroup() %>% 
  dplyr::arrange(desc(num))
```

```{r}
df_cancer_stage$Surgery.Decision <- revalue(df_cancer_stage$Surgery.Decision, 
c("Recommended, unknown if performed"= "Recommended",
  
"Recommended but not performed, patient refused"= "Recom & refused",

"Recommended but not performed, unknown reason"= "Recom & not performed",

"Not recommended, contraindicated due to other cond; autopsy only (1973-2002)"= "Not recom & contraindicated",

"Not performed, patient died prior to recommended surgery"= "Not performed & died",

"Unknown; death certificate; or autopsy only (2003+)"= "Unknown"))
```

  The plot below shows that, even given recommendation, some patients with oral cavity cancer in nearly all the stages, even the early stage, have given up the surgery. 
  Combining the infomation given by the status of chemo and radiation therapy, we would like to choose oral cavity cancer as our research direction.

```{r}
# Top 10 Giving-Up-Surgery Cancer
df_cancer_stage %>% 
  dplyr::filter(Surgery.Decision %in% c("Recommended","Recom & refused","Recom & not performed")) %>% 
  mutate(percent = num/sum(num)) %>% 
  mutate(Cancer_Stage = reorder(Cancer_Stage,percent)) %>% 
  top_n(10,wt = percent) %>% 
  ggplot() + 
  geom_bar(aes(x = Cancer_Stage,y = percent, fill = Surgery.Decision),stat = "identity") + 
  labs(title = "Top 10 Cancers that are Most Likely to Giving Up Surgery", 
       x = "Cancer with Stage",y = "Percentage",
       fill = "Surgery Decision") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c("orange", "royalblue","limegreen"))+
  coord_flip()
```


```{r message=FALSE, warning=FALSE}
# Chemo
df_cancer_chemo <- DF %>% 
  dplyr::select(Site,AJCC.7.Stage,Chemotherapy,Study.ID) %>% 
  unite("Cancer_Stage",Site,AJCC.7.Stage,sep = "-") %>% 
  dplyr::group_by(Cancer_Stage,Chemotherapy) %>% 
  dplyr::summarise(num = n()) %>% ungroup() %>% 
  dplyr::arrange(desc(num))
```

```{r}
df_cancer_chemo %>% 
  mutate(percent = num/sum(num)) %>% 
  mutate(Cancer_Stage = reorder(Cancer_Stage,percent)) %>% 
  top_n(10,wt = percent) %>% 
  ggplot() + 
  geom_bar(aes(x = Cancer_Stage,y = percent, fill = Chemotherapy),stat = "identity") + 
  labs(title = "Top 10 Cancers by Chemo Status", 
       x = "Cancer with Stage",y = "Percentage",
       fill = "Chemo Status") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c("#003399","#00CC33"))+
  coord_flip()
```

```{r message=FALSE, warning=FALSE}
# Radiation
df_cancer_radi <- DF %>% 
  dplyr::select(Site,AJCC.7.Stage,Radiation,Study.ID) %>% 
  unite("Cancer_Stage",Site,AJCC.7.Stage,sep = "-") %>% 
  dplyr::group_by(Cancer_Stage,Radiation) %>% 
  dplyr::summarise(num = n()) %>% ungroup() %>% 
  dplyr::arrange(desc(num))
```

```{r}
df_cancer_radi %>% 
  mutate(percent = num/sum(num)) %>% 
  mutate(Cancer_Stage = reorder(Cancer_Stage,percent)) %>% 
  top_n(10,wt = percent) %>% 
  ggplot() + 
  geom_bar(aes(x = Cancer_Stage,y = percent, fill = Radiation),stat = "identity") + 
  labs(title = "Top 10 Cancers by Radiation Status", 
       x = "Cancer with Stage",y = "Percentage",
       fill = "Radiation Status") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c("#99CCCC","#6633FF"))+
  theme(legend.position = "bottom", legend.direction = "horizontal")+
  coord_flip()
```



```{r echo=FALSE}
# Data of Oral Cavity
df_oral <- DF %>% dplyr::filter(Site == "Oral Cavity")
```

## Features vs `Surgery Performed` 

  To begin with, we would like to take a look at the condition of surgery implementation.

**State**

### Environment Attributes by State

  Household Income, Poverty, Unemployment, Language Isolation and Education are all recorded on county level, which can also be regarded as important attributes of the respondents' `SEER.Registry`, so we would like to explore their distribution by states.
  
  For Poverty, the state of Georgia has the highest poverty rate, over 15% in average. For Unemployment, the unemployment rates are quite similar across those 4 states. For Household Income, the state of Georgia has the lowest household income in average. For Language Isolation, the state of Alaska has the lowest language isolation rate, while the state of California has the highest.
  

```{r message=FALSE}
g1 <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(X..Persons.Below.Poverty)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Below Poverty Percentage vs State", x = "State", y = "Below Poverty Percentage")

g2 <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(X..Unemployed.ACS.2013.2017)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Unemployed Percentage vs State", x = "State", y = "Unemployed Percentage")

g3 <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(Median.Household.Income)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Median Household Income vs State", x = "State", y = "Median Household Income")

g4 <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(X..Language.isolation.ACS.2013.2017..households.)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Language Isolation Percentage vs State", x = "State", y = "Language Isolation Percentage")
```

```{r}
gridExtra::grid.arrange(g1, g2, g3, g4, nrow = 2)
```

```{r message=FALSE}
edu_9 <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Avg = mean(X...9th.Grade.Education))
edu_h <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Avg = mean(X...High.School.Education))
edu_b <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Avg = mean(X...Bachelors.Education))
edu_9$Avg1 <- edu_h$Avg
edu_9$Avg2 <- edu_b$Avg

df_int <- stack(edu_9[,2:4])
df_state_edu <- data.frame(rep(c("Alaska", "California", "Connecticut", "Georgia"), 3), df_int)
colnames(df_state_edu) <- c("State", "values", "education")

state <- c(rep("Alaska", 3), rep("California", 3), rep("Connecticut", 3), rep("Georgia", 3))
edu <- rep(c("9th Grade", "High School", "Bachelor"), 4)
values <- c(2.66, 7.61, 28.99, 9.242076, 16.874751, 32.903706, 4.116577, 9.667285, 38.037514, 5.239321, 14.709081, 27.351167)
```

  For Education, the state of California has the highest proportion of respondents are educated below 9th Grade and the highest proportion of repondents are educated below High School. Meanwhile, the state of Connecticut has the highest proportion of respondents are educated below Bechelor.
  
```{r message=FALSE}
df_state_edu2 <- data.frame(state, edu, values)
df_state_edu2 %>% 
  ggplot(aes(fill = edu, x = state, y = values/100)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(title = "Education Status per State", x = "State", y = "Percentage", fill = "Education") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))
  
```

  By examining surgery implementation condition by state, we found that California has the largest population of patients researched in the data. While Alaska and Connecticut have the largest proportion of patients who have performed the surgery.
  
```{r message=FALSE}
# Registry State vs. Whether the surgery implemented
p1 <- ggplot(df_oral, aes(x = SEER.Registry, fill = Surgery.Performed.) ) +  
    geom_bar( aes(y = ..count../sum(..count..) ) ) +
    labs(title = "Surgery Implementation by State",
         x = "Registered State",
         y = "Percentage") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred") ) 

p2 <- df_oral %>% select(SEER.Registry, Study.ID, Surgery.Performed.) %>% 
  dplyr::group_by(SEER.Registry, Surgery.Performed.) %>% 
  dplyr::summarise(num = n()) %>% 
  ungroup() %>% 
  dplyr::mutate(percent = num/sum(num)) %>% 
  
  ggplot()+
  geom_bar(aes(x = SEER.Registry, y = percent, fill = Surgery.Performed.), position = "fill", stat = "identity") + 
  labs(
         x = "Registered State",
         y = "Percentage") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred") )

gridExtra::grid.arrange(p1,p2)
```

**Race & Gender**

  Before looking at the surgery condition by race and gender, we would like to take a look at the distribution of race by gender. The plot above shows that, more than 60% of observations are white people, and more than 40% of observations are white male, which indicates that there would be an imbalance problem in the data.
    
```{r message=FALSE, warning=FALSE}
# Distribution of race by gender
df_oral %>% select(Race,Sex,Study.ID) %>% 
    dplyr::group_by(Race,Sex) %>% 
    dplyr::summarise(num = n()) %>% 
    ungroup() %>% 
    dplyr::mutate(percent = num/sum(num)) %>% 
    
    ggplot() +
    geom_bar(aes(x = Race,y = percent,fill = Sex),stat = "identity") +
    labs(title = "Distribution of Race by Gender",x = "Race", y = "Percentage") + 
    scale_fill_manual("Gender", values = c("pink", "skyblue"))+
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 10, hjust = 0.5, vjust = 0.5))
```
    
  For surgery condition, the plot below shows that Asian of Pacific Islander has the largest proportion of performing the surgery while the Black has the smallest one. Female and Male has almost the same proportion to accept the surgery, which seems that there is no gender biases.
    
```{r message=FALSE}
# Race vs. Whether the surgery implemented

p3 <- df_oral %>% select(Race, Study.ID, Surgery.Performed.) %>% 
  dplyr::group_by(Race, Surgery.Performed.) %>% 
  dplyr::summarise(num = n()) %>% 
  ungroup() %>% 
  dplyr::mutate(percent = num/sum(num)) %>% 
  
  ggplot()+
  geom_bar(aes(x = Race, y = percent, fill = Surgery.Performed.), position = "fill", stat = "identity") + 
  labs(title = "Surgery Implementation by Race & Gender",
         x = "Race",
         y = "Percentage") +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 10, hjust = 0.5, vjust = 0.5)) + 
    guides(fill = F)

# Gender vs. Whether the surgery implemented

p4 <- df_oral %>% select(Sex, Study.ID, Surgery.Performed.) %>% 
  dplyr::group_by(Sex, Surgery.Performed.) %>% 
  dplyr::summarise(num = n()) %>% 
  ungroup() %>% 
  dplyr::mutate(percent = num/sum(num)) %>% 
  
  ggplot()+
  geom_bar(aes(x = Sex, y = percent, fill = Surgery.Performed.), position = "fill", stat = "identity") + 
  labs(
         x = "Gender",
         y = "Percentage") +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))

gridExtra::grid.arrange(p3,p4)
```

  After looking at distribution by gender and race separately, we would like to put these 2 factors together to explore the distribution of performed surgery.
  
  Then, we find that the white male has the largest proportion of performing a surgery. 
    
```{r}
# Gender cross Race  vs.  Whether the surgery implemented
df2 <- df_oral[,c("Sex","Race", "Surgery.Performed.")] 
df2$Sex <- revalue(df2$Sex, c("Female"="F"))
df2$Sex <- revalue(df2$Sex,c("Male"="M"))
df2$Race <- revalue(df2$Race,c("Asian or Pacific Islander"="Asian/PI"))
df2$Race <- revalue(df2$Race,c("American Indian/Alaska Native"="Indigen"))
df2 <-  unite(df2, "Sex_Race", Sex, Race, sep = "_", remove = FALSE)
```


```{r}
ggplot(df2, aes(x = Sex_Race, fill = Surgery.Performed.) ) +  
    geom_bar( aes(y = ..count../sum(..count..) ) ) +
    labs(title = "Surgery Implementation by Gender_Race",
         x = "Gender_Race",
         y = "Percentage")+
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) + coord_flip()
```
**Age**

  The age feature is recorded as a continuous variable, so we make a density plot instead of the bar plot. In the below plot, we find that the distributions of performing surgery and not performing surgery on ages are mostly overlapping. However, we can still see that, for people older than 75 years old, they are more likely to perform the surgery; while for people younger than 75 years old, they are more likely to not perform the surgery.
 
```{r message=FALSE}
ggplot(data=df_oral, aes(x=Age.at.Diagnosis, group=Surgery.Performed., fill=Surgery.Performed.)) +
    geom_density(adjust=1.5, alpha=.4) + 
  labs(title = "Surgery Performed Status by Age",
       x = "Age at Diagnosis",y = "Density",
       fill = "Surgery Performed")
```

**Insurance**

  Besides of the demographic features, we also concern about the impact of social biases like insurance on whether to perform the surgery or not. From below chart, we can see that the large majority of repondents are insured and insured people are most likely to perform the surgery.

```{r message=FALSE, warning=FALSE}
p5 <- df_oral %>% select(Insurance,Surgery.Performed.,Study.ID) %>% 
    dplyr::group_by(Insurance,Surgery.Performed.) %>% 
    dplyr::summarise(num = n()) %>% 
    ungroup() %>% 
    mutate(percent = num/sum(num)) %>% 
    ggplot() +
    geom_bar(aes(x = Insurance,y = percent,fill = Surgery.Performed.),stat = "identity") +
    labs(title = "Surgery Implementation by Insurance Type",x = NULL, y = "Percentage") + 
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) + coord_flip()

p6 <- df_oral %>% select(Insurance,Surgery.Performed.,Study.ID) %>% 
    dplyr::group_by(Insurance,Surgery.Performed.) %>% 
    dplyr::summarise(num = n()) %>% 
    ungroup() %>% 
    mutate(percent = num/sum(num)) %>% 
    ggplot() +
    geom_bar(aes(x = Insurance,y = percent,fill = Surgery.Performed.), position = "fill", stat = "identity") +
    labs(x = NULL, y = "Percentage") + 
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) + coord_flip()

gridExtra::grid.arrange(p5,p6)
```

## Trend

  After exploring the relationship between all kinds of features and the surgery implementation situation, we want to research the change of surgery implementation on time series.
  
  From the below plot, for the trend of amount of performed surgery, we can see that as time goes by, there is an upper trend. 
    
```{r message=FALSE, warning=FALSE}
df_oral %>% 
    select(Year.of.Diagnosis,Surgery.Performed.,Study.ID) %>% 
    mutate(Year.of.Diagnosis = as.character(Year.of.Diagnosis)) %>% 
    dplyr::group_by(Year.of.Diagnosis,Surgery.Performed.) %>% 
    dplyr::summarise(num = n()) %>% ungroup() %>% 
    ggplot() + 
    geom_line(aes(x = Year.of.Diagnosis,
                  y = num,
                  group = Surgery.Performed.,
                  col = Surgery.Performed.),size = 1) +
    labs(title = "Performed Surgery Trend by Year",
         x = "Year of Dignosis",y = "Count") +
    scale_color_manual("Surgery Implementation",values = c("grey55", "darkred"))
```


# Feature Engineering

## Feature Establishment

  The columns `Surgery.Performed.`,`Radiation` and `Chemotherapy` contain information of the actual therapy that the patients have taken. It would be more informative to combine these 3 columns together to reflect patients' treatment as a whole. Hence, we create a column `True_Therapy` to get the aggregated information.
  
```{r message=FALSE, warning=FALSE}
DF_pre <- df_oral

# Combine variables of Surgery Performed, Chemotherapy and Radiation into one varibale called True_Therapy

DF_pre$rad = ifelse(DF_pre$Radiation =="No radiation and/or cancer-directed surgery", "No", "Yes")
DF_pre$True_Therapy = rep(NA,nrow(DF_pre))

for (i in 1:nrow(DF_pre)){
  
  if(DF_pre$Surgery.Performed.[i] == "Yes" && DF_pre$Chemotherapy[i] == "Yes" && DF_pre$rad[i] == "Yes"){
    DF_pre$True_Therapy[i] = "S+C+R"
  }
  else if(DF_pre$Surgery.Performed.[i] == "Yes" && DF_pre$Chemotherapy[i] == "Yes" && DF_pre$rad[i] == "No"){
    DF_pre$True_Therapy[i] = "S+C"
  }
  else if(DF_pre$Surgery.Performed.[i] == "Yes" && DF_pre$Chemotherapy[i] == "No/Unknown" && DF_pre$rad[i] == "Yes"){
    DF_pre$True_Therapy[i] = "S+R"
  }
  else if(DF_pre$Surgery.Performed.[i] == "Yes" && DF_pre$Chemotherapy[i] == "No/Unknown" && DF_pre$rad[i] == "No"){
    DF_pre$True_Therapy[i] = "S"
  }
  else if(DF_pre$Surgery.Performed.[i] == "No" && DF_pre$Chemotherapy[i] == "No/Unknown" && DF_pre$rad[i] == "No"){
    DF_pre$True_Therapy[i] = "No treatment"
  }
  else if(DF_pre$Surgery.Performed.[i] == "No" && DF_pre$Chemotherapy[i] == "No/Unknown" && DF_pre$rad[i] == "Yes"){
    DF_pre$True_Therapy[i] = "R"
  }
  else if(DF_pre$Surgery.Performed.[i] == "No" && DF_pre$Chemotherapy[i] == "Yes" && DF_pre$rad[i] == "No"){
    DF_pre$True_Therapy[i] = "C"
  }
  else if(DF_pre$Surgery.Performed.[i] == "No" && DF_pre$Chemotherapy[i] == "Yes" && DF_pre$rad[i] == "Yes"){
    DF_pre$True_Therapy[i] = "C+R"
  }
  
}
```

  Meanwhile, the 2021 version **NCCN Guideline** has also provided the information of what kinds of treatments patients with oral cavity cancer are supposed to take. 
  
  Since the recommended treatment is based on `AJCC.7.Stage`, where contains information of tumor size (`Size`) and lymph nodes status (`Lymph.Nodes`), we will create a column `Rec_Therapy` indicating the treatments that a patient is supposed to have under **NCCN Guideline**.
  
```{r message=FALSE, warning=FALSE}
# Create the canser stage for oral cavity
df_recommend <- DF_pre %>% 
  select(Site,Size,AJCC.7.Stage,Study.ID,True_Therapy) %>%
  dplyr::filter(Site=="Oral Cavity") %>%
  unite("Cancer_Stage",Site,AJCC.7.Stage,sep = "-") %>% 
  mutate(Cancer_Stage = factor(Cancer_Stage,levels = c("Oral Cavity-I","Oral Cavity-II","Oral Cavity-III","Oral Cavity-IVA","Oral Cavity-IVB","Oral Cavity-IVC","Oral Cavity-IVNOS")))
```

```{r message=FALSE, warning=FALSE}
# Create the variable Rec_Therapy that shows the therapy combinations for patients based on the NCCN guideline.
df_recommend$Rec_Therapy <- rep(NA,nrow(df_recommend))
for(i in 1:nrow(df_recommend)){
  if(df_recommend$Cancer_Stage[i]=="Oral Cavity-I"){
    df_recommend$Rec_Therapy[i]="S+R"
  }
  else if(df_recommend$Cancer_Stage[i]=="Oral Cavity-II"){
    df_recommend$Rec_Therapy[i]="S+R"
  }
  else{
    df_recommend$Rec_Therapy[i]="S"
  }
}
```

  For equity issue, we believe it would be more important to examine the matchness of recommended therapies and the real therapies that patients actually get. 
  
  So, we create a column `Match` to evaluate such matchness. `Match` is a binary feature, it equals to 1 when `True_Therapy` and `Rec_Therapy` are completely matched or `Rec_Therapy` is the subset of `True_Therapy`(eg. A patient is supposed to take surgery based on NCCN Guideline, and this patient actully has taken surgery and chemo), and it equals to 0 otherwise.
  
```{r message=FALSE, warning=FALSE}
# Create the variable Match with binnary outcomes.
df_recommend$Match = ifelse(df_recommend$True_Therapy == df_recommend$Rec_Therapy, 1,0)
for(i in 1:nrow(df_recommend)){
  if(df_recommend$True_Therapy[i]=="S+R" && df_recommend$Rec_Therapy[i] == "S"){
    df_recommend$Match[i] = 1
  }
  else if(df_recommend$True_Therapy[i]=="S+C+R" && df_recommend$Rec_Therapy[i] == "S"){
    df_recommend$Match[i] = 1
  }
  else if(df_recommend$True_Therapy[i]=="S+C+R" && df_recommend$Rec_Therapy[i] == "S+R"){
    df_recommend$Match[i] = 1
  }
}
```


```{r}
# Merger the data for modeling
oral_cancer <- left_join(DF_pre[,-which((names(DF_pre) %in% c("Site","Subsite","AJCC.7.Stage","Size","Lymph.Nodes","Mets","Surgery.Performed.","Surgery.Decision","Radiation","Chemotherapy","rad","True_Therapy")))],df_recommend,by = "Study.ID")

oral_cancer <- 
  oral_cancer %>% dplyr::rename(
  `9th.Education` = X...9th.Grade.Education,
  HS.Education = X...High.School.Education,
  Bachelors.Education = X...Bachelors.Education,
  Below.Poverty = X..Persons.Below.Poverty,
  Unemployed = X..Unemployed.ACS.2013.2017,
  Language.Isolation = X..Language.isolation.ACS.2013.2017..households.)

# Data set for modeling
# write.csv(oral_cancer,"oral_cancer_data.csv",row.names = F)
```


# Modeling

```{r}

```


