---
title: "679 Final Project"
author: "Team 9"
date: "4/12/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(plyr)
library(reshape2)
library(tidyr)
library(viridis)
library(psych)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
DF <- read.csv("transformed.csv", encoding="UTF-8")
describe(DF)
```

```{r}
colnames(DF) <- c("Study.ID","Sex",                                    "Year.of.Diagnosis","Age.at.Diagnosis","Race",                                       "Insurance","SEER.Registry","X...9th.Grade.Education",                         "X...High.School.Education","X...Bachelors.Education",                         "X..Persons.Below.Poverty","X..Unemployed.ACS.2013.2017",                     "Median.Household.Income",                      "X..Language.isolation.ACS.2013.2017..households.","Site",                              "Subsite","AJCC.7.Stage","Size",                                            "Lymph.Nodes","Mets","Cause.of.Death",                                  "Surgery.Performed.","Surgery.Decision","Radiation",                                    "Chemotherapy")
colnames(DF)
```

# EDA

## `Surgery Decision` vs `Surgery Implementation`

In the categories of surgery decision, we can see that the top 2 frequent categories are *Surgery performed* and *Not recommended*. Besides, only the patients whose surgery decision is "surgery performed" actually took the surgery.

```{r message=FALSE}
df_surgery <- DF[,c("Surgery.Decision","Sex", "Race", "Insurance","SEER.Registry", "Surgery.Performed.", "Study.ID")] 
df_surgery$Surgery.Decision <- revalue(df_surgery$Surgery.Decision,
c("Recommended, unknown if performed"= "Recommended",
  "Recommended but not performed, patient refused"= "Recom & refused",
  "Recommended but not performed, unknown reason"= "Recom & not performed",
  "Not recommended, contraindicated due to other cond; autopsy only (1973-2002)"= "Not recom & contraindicated",
  "Not performed, patient died prior to recommended surgery"= "Not performed & died",
  "Unknown; death certificate; or autopsy only (2003+)"= "Unknown"))


df_surgery$Race <- revalue(df_surgery$Race,
c("Asian or Pacific Islander"="Asian/PI",
  "American Indian/Alaska Native"="Indigen"))
```


```{r message=FALSE}
# Surgery Implementation by Surgery Decision
df_surgery %>% select(Surgery.Decision,Surgery.Performed.,Study.ID) %>% 
    dplyr::group_by(Surgery.Decision,Surgery.Performed.) %>% 
    dplyr::summarise(num = n()) %>% 
    ungroup() %>% 
    mutate(percent = num/sum(num)) %>% 
    ggplot() +
    geom_bar(aes(x = Surgery.Decision,y = percent,fill = Surgery.Performed.),stat = "identity") +
    labs(title = "Surgery Implementation by Surgery Decision",x = NULL, y = "Percentage") + 
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) + coord_flip()
```

# Environment Attributes by State

  Household Income, Poverty, Unemployement, Language Isolation and Education are important attributes of the respondents' `SEER.Registry`, so we would like to explore their distribution by states.
  
  For Poverty, the state of Georgia has the highest proverty rate, over 15% in average. For Unemployment, the unemployment rates are quite similar across those 4 states. For Household Income, the state of Georgia has the lowest household income in average. For Language Isolation, the state of Alaska has the lowest language isolation rate, while the state of California has the highest.
  

```{r message=FALSE}
g1 <- DF %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(X..Persons.Below.Poverty)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Below Poverty Percentage vs State", x = "State", y = "Below Poverty Percentage")

g2 <- DF %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(X..Unemployed.ACS.2013.2017)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Unemployed Percentage vs State", x = "State", y = "Unemployed Percentage")

g3 <- DF %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(Median.Household.Income)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Median Household Income vs State", x = "State", y = "Median Household Income")

g4 <- DF %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(X..Language.isolation.ACS.2013.2017..households.)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Language Isolation Percentage vs State", x = "State", y = "Language Isolation Percentage")
```

```{r}
gridExtra::grid.arrange(g1, g2, g3, g4, nrow = 2)
```

```{r message=FALSE}
edu_9 <- DF %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Avg = mean(X...9th.Grade.Education))
edu_h <- DF %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Avg = mean(X...High.School.Education))
edu_b <- DF %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Avg = mean(X...Bachelors.Education))
edu_9$Avg1 <- edu_h$Avg
edu_9$Avg2 <- edu_b$Avg

df_int <- stack(edu_9[,2:4])
df_state_edu <- data.frame(rep(c("Alaska", "California", "Connecticut", "Georgia"), 3), df_int)
colnames(df_state_edu) <- c("State", "values", "education")

state <- c(rep("Alaska", 3), rep("California", 3), rep("Connecticut", 3), rep("Georgia", 3))
edu <- rep(c("9th Grade", "High School", "Bachelor"), 4)
values <- c(df_state_edu$values[1], df_state_edu$values[5], df_state_edu$values[9], df_state_edu$values[2], df_state_edu$values[6], df_state_edu$values[10], df_state_edu$values[3], df_state_edu$values[7], df_state_edu$values[11], df_state_edu$values[4], df_state_edu$values[8], df$values[12])
```


  For Education, the state of California has the highest proportion of respondents are educated below 9th Grade and the highest proportion of repondents are educated below High School. Meanwhile, the state of Connecticut has the highest proportion of respondents are educated below Bechelor.
  
```{r message=FALSE}
df_state_edu2 <- data.frame(state, edu, values)
df_state_edu2 %>% 
  ggplot(aes(fill = edu, x = state, y = values/100)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(title = "Education Status per State", x = "State", y = "Percentage", fill = "Education") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))
  
```


## Feature: `Surgery Performed` 

  To begin with, we would like to take a look at the condition of surgery implementation.

**State**

  By examining surgery implemetation condition by state, we found that California has the largest percentage of patient to take the surgery among the 4 states.
  
```{r message=FALSE}
# Registry State vs. Whether the surgery implemented
p1 <- ggplot(DF, aes(x = SEER.Registry, fill = Surgery.Performed.) ) +  
    geom_bar( aes(y = ..count../sum(..count..) ) ) +
    labs(title = "Surgery Implementation by State",
         x = "Registered State",
         y = "Percentage") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred") ) 

p2 <- DF %>% select(SEER.Registry, Study.ID, Surgery.Performed.) %>% 
  dplyr::group_by(SEER.Registry, Surgery.Performed.) %>% 
  dplyr::summarise(num = n()) %>% 
  ungroup() %>% 
  dplyr::mutate(percent = num/sum(num)) %>% 
  
  ggplot()+
  geom_bar(aes(x = SEER.Registry, y = percent, fill = Surgery.Performed.), position = "fill", stat = "identity") + 
  labs(
         x = "Registered State",
         y = "Percentage") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred") )

gridExtra::grid.arrange(p1,p2)
```

**Race & Gender**
  Before looking at the surgery condition by race and gender, we would like to take a look at the distribution of race by gender. The plot above shows that, more than 60% of observations are white people, and more than 50% of observations are white male, which indicates that there would be an imbalance problem in the data.
    
```{r message=FALSE, warning=FALSE}
# Distribution of race by gender
DF %>% select(Race,Sex,Study.ID) %>% 
    dplyr::group_by(Race,Sex) %>% 
    dplyr::summarise(num = n()) %>% 
    ungroup() %>% 
    dplyr::mutate(percent = num/sum(num)) %>% 
    
    ggplot() +
    geom_bar(aes(x = Race,y = percent,fill = Sex),stat = "identity") +
    labs(title = "Distribution of Race by Gender",x = "Race", y = "Percentage") + 
    scale_fill_manual("Gender", values = c("pink", "skyblue"))+
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 10, hjust = 0.5, vjust = 0.5))
```
    
  For surgery condition, the plot below shows that the proportion of white people who take the surgery and male who take he surgery is the largest.
    
```{r message=FALSE}
# Race vs. Whether the surgery implemented

p3 <- DF %>% select(Race, Study.ID, Surgery.Performed.) %>% 
  dplyr::group_by(Race, Surgery.Performed.) %>% 
  dplyr::summarise(num = n()) %>% 
  ungroup() %>% 
  dplyr::mutate(percent = num/sum(num)) %>% 
  
  ggplot()+
  geom_bar(aes(x = Race, y = percent, fill = Surgery.Performed.), position = "fill", stat = "identity") + 
  labs(title = "Surgery Implementation by Race & Gender",
         x = "Race",
         y = "Percentage") +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 10, hjust = 0.5, vjust = 0.5)) + 
    guides(fill = F)

# Gender vs. Whether the surgery implemented

p4 <- DF %>% select(Sex, Study.ID, Surgery.Performed.) %>% 
  dplyr::group_by(Sex, Surgery.Performed.) %>% 
  dplyr::summarise(num = n()) %>% 
  ungroup() %>% 
  dplyr::mutate(percent = num/sum(num)) %>% 
  
  ggplot()+
  geom_bar(aes(x = Sex, y = percent, fill = Surgery.Performed.), position = "fill", stat = "identity") + 
  labs(
         x = "Gender",
         y = "Percentage") +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```

```{r}
gridExtra::grid.arrange(p3,p4)
```

  After looking at distribution by gender and race separately, we would like to put these 2 factors together to explore the distribution of performed surgery.
  
  Then, we find that the white male has the largest proportion of performing a surgery. 
    
```{r}
# Gender cross Race  vs.  Whether the surgery implemented
DF2 <- DF[,c("Sex","Race", "Surgery.Performed.")] 
DF2$Sex <- revalue(DF2$Sex, c("Female"="F"))
DF2$Sex <- revalue(DF2$Sex,c("Male"="M"))
DF2$Race <- revalue(DF2$Race,c("Asian or Pacific Islander"="Asian/PI"))
DF2$Race <- revalue(DF2$Race,c("American Indian/Alaska Native"="Indigen"))
DF2 <-  unite(DF2, "Sex_Race", Sex, Race, sep = "_", remove = FALSE)
```


```{r}
ggplot(DF2, aes(x = Sex_Race, fill = Surgery.Performed.) ) +  
    geom_bar( aes(y = ..count../sum(..count..) ) ) +
    labs(title = "Surgery Implementation by Gender_Race",
         x = "Gender_Race",
         y = "Percentage")+
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) + coord_flip()
```

**Insurance**
```{r message=FALSE, warning=FALSE}
p5 <- DF %>% select(Insurance,Surgery.Performed.,Study.ID) %>% 
    dplyr::group_by(Insurance,Surgery.Performed.) %>% 
    dplyr::summarise(num = n()) %>% 
    ungroup() %>% 
    mutate(percent = num/sum(num)) %>% 
    ggplot() +
    geom_bar(aes(x = Insurance,y = percent,fill = Surgery.Performed.),stat = "identity") +
    labs(title = "Surgery Implementation by Insurance Type",x = NULL, y = "Percentage") + 
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) + coord_flip()

p6 <- DF %>% select(Insurance,Surgery.Performed.,Study.ID) %>% 
    dplyr::group_by(Insurance,Surgery.Performed.) %>% 
    dplyr::summarise(num = n()) %>% 
    ungroup() %>% 
    mutate(percent = num/sum(num)) %>% 
    ggplot() +
    geom_bar(aes(x = Insurance,y = percent,fill = Surgery.Performed.), position = "fill", stat = "identity") +
    labs(x = NULL, y = "Percentage") + 
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) + coord_flip()

```

```{r}
gridExtra::grid.arrange(p5,p6)
```

**Trend**
    For the trend of amount of performed surgery, we can see that as time goes by, there is an upper trend. 
```{r message=FALSE, warning=FALSE}
DF %>% 
    select(Year.of.Diagnosis,Surgery.Performed.,Study.ID) %>% 
    mutate(Year.of.Diagnosis = as.character(Year.of.Diagnosis)) %>% 
    dplyr::group_by(Year.of.Diagnosis,Surgery.Performed.) %>% 
    dplyr::summarise(num = n()) %>% ungroup() %>% 
    ggplot() + 
    geom_line(aes(x = Year.of.Diagnosis,
                  y = num,
                  group = Surgery.Performed.,
                  col = Surgery.Performed.),size = 1) +
    labs(title = "Performed Surgery Trend by Year",
         x = "Year of Dignosis",y = "Count") +
    scale_color_manual("Surgery Implementation",values = c("grey55", "darkred"))
```

```{r message=FALSE, warning=FALSE}
ggplot(data=DF, aes(x=Age.at.Diagnosis, group=Surgery.Performed., fill=Surgery.Performed.)) +
    geom_density(adjust=1.5, alpha=.4) + 
  labs(title = "Surgery Performed Status by Age",
       x = "Age at Diagnosis",y = "Density",
       fill = "Surgery Performed")
```

```{r message=FALSE, warning=FALSE}
ggplot(data=DF, aes(x=Size, group=Surgery.Performed., fill=Surgery.Performed.)) +
    geom_density(adjust=1.5, alpha=.4) + 
  labs(title = "Surgery Performed Status by Tumor Size",
       x = "Tumor Size (mm)",y = "Density",
       fill = "Surgery Performed")
```

## Feature: `Surgery Decision` 

```{r message=FALSE, warning=FALSE}
df_cancer_stage <- DF %>% 
  dplyr::select(Site,AJCC.7.Stage,Surgery.Decision,Study.ID) %>% 
  unite("Cancer_Stage",Site,AJCC.7.Stage,sep = "-") %>% 
  dplyr::group_by(Cancer_Stage,Surgery.Decision) %>% 
  dplyr::summarise(num = n()) %>% ungroup() %>% 
  dplyr::arrange(desc(num))
```


```{r}
df_cancer_stage$Surgery.Decision <- revalue(df_cancer_stage$Surgery.Decision, 
c("Recommended, unknown if performed"= "Recommended",
  
"Recommended but not performed, patient refused"= "Recom & refused",

"Recommended but not performed, unknown reason"= "Recom & not performed",

"Not recommended, contraindicated due to other cond; autopsy only (1973-2002)"= "Not recom & contraindicated",

"Not performed, patient died prior to recommended surgery"= "Not performed & died",

"Unknown; death certificate; or autopsy only (2003+)"= "Unknown"))
```

  The plot below shows that, even given recommendation, more than 20% of patients with cancer in Oropharynx with stage IVA are likely to give up surgery.

```{r}
# Top 10 Giving-Up-Surgery Cancer
df_cancer_stage %>% 
  dplyr::filter(Surgery.Decision %in% c("Recommended","Recom & refused","Recom & not performed")) %>% 
  mutate(percent = num/sum(num)) %>% 
  mutate(Cancer_Stage = reorder(Cancer_Stage,percent)) %>% 
  top_n(10,wt = percent) %>% 
  ggplot() + 
  geom_bar(aes(x = Cancer_Stage,y = percent, fill = Surgery.Decision),stat = "identity") + 
  labs(title = "Top 10 Cancers that are Most Likely to Giving Up Surgery", 
       x = "Cancer with Stage",y = "Percentage",
       fill = "Surgery Decision") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c("orange", "royalblue","limegreen"))+
  coord_flip()
```



