---
title: "679 Final Project"
author: "Team 9(Chenghao Meng, Chun Gao, Donghao Xue, Jingwen Xu)"
date: "5/5/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(plyr)
library(reshape2)
library(tidyr)
library(viridis)
library(psych)
library(rstanarm)
library(arm)
library(randomForest)
library(caret)
library(e1071)
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

  The treatment and management of patients with chronic and severe illnesses have been affected by several different aspects of factors, not just their medical care condition but their social background. From a practical and data acquisition point of view, the data sets provided by our clients mainly focus on cancer research.

  Our main goal of the project is to explore how biases (race, gender, education level and other factors including) will affect the matching between the treatments that a patient should have and actually have. Besides, to simplify our project we would like to focus on oral cavity cancer and set particular individuals as a reference group.

# Data Description

  We have a dataset with 23,291 rows and 25 columns. This dataset contains information of patients who have cancers in 7 different sites, including Oral Cavity, Sinonasal, Larynx, Salivary Gland, Oropharynx, Hypopharynx and Nasopharynx. The columns are showed as below:

```{r echo=FALSE}
DF <- read.csv("transformed.csv", encoding="UTF-8")
```

```{r echo=FALSE}
colnames(DF) <- c("Study.ID","Sex",                                    "Year.of.Diagnosis","Age.at.Diagnosis","Race",                                       "Insurance","SEER.Registry","X...9th.Grade.Education",                         "X...High.School.Education","X...Bachelors.Education",                         "X..Persons.Below.Poverty","X..Unemployed.ACS.2013.2017",                     "Median.Household.Income",                      "X..Language.isolation.ACS.2013.2017..households.","Site",                              "Subsite","AJCC.7.Stage","Size",                                            "Lymph.Nodes","Mets","Cause.of.Death",                                  "Surgery.Performed.","Surgery.Decision","Radiation",                                    "Chemotherapy")
colnames(DF)
```

# Exploratory Data Analysis

## Cancer Type Selection

  There are seven kinds of cancer with different stages in this dataset. To simplify our research, we decide to choose only one kind of cancer to do the analysis.

  So, we firstly explore the top 10 cancers (including the stages) that the patients are most likely to give up the surgery. Giving up the surgery can suggest that the patients are affected by some factors such as biases in the society.

  The plot below shows that, even given recommendation, some patients with oral cavity cancer in nearly all the stages, even the early stage, have given up the surgery. Combining the information given by the status of chemo and radiation therapy, we would like to choose oral cavity cancer as our research direction.

```{r message=FALSE, echo=FALSE}
# Surgery.Decision
df_cancer_stage <- DF %>% 
  dplyr::select(Site,AJCC.7.Stage,Surgery.Decision,Study.ID) %>% 
  unite("Cancer_Stage",Site,AJCC.7.Stage,sep = "-") %>% 
  dplyr::group_by(Cancer_Stage,Surgery.Decision) %>% 
  dplyr::summarise(num = n()) %>% ungroup() %>% 
  dplyr::arrange(desc(num))
```

```{r echo=FALSE}
df_cancer_stage$Surgery.Decision <- revalue(df_cancer_stage$Surgery.Decision, 
c("Recommended, unknown if performed"= "Recommended",
  
"Recommended but not performed, patient refused"= "Recom & refused",

"Recommended but not performed, unknown reason"= "Recom & not performed",

"Not recommended, contraindicated due to other cond; autopsy only (1973-2002)"= "Not recom & contraindicated",

"Not performed, patient died prior to recommended surgery"= "Not performed & died",

"Unknown; death certificate; or autopsy only (2003+)"= "Unknown"))
```

```{r echo=FALSE}
# Top 10 Giving-Up-Surgery Cancer
df_cancer_stage %>% 
  dplyr::filter(Surgery.Decision %in% c("Recommended","Recom & refused","Recom & not performed")) %>% 
  mutate(percent = num/sum(num)) %>% 
  mutate(Cancer_Stage = reorder(Cancer_Stage,percent)) %>% 
  top_n(10,wt = percent) %>% 
  ggplot() + 
  geom_bar(aes(x = Cancer_Stage,y = percent, fill = Surgery.Decision),stat = "identity") + 
  labs(title = "Top 10 Cancers that are Most Likely to Giving Up Surgery", 
       x = "Cancer with Stage",y = "Percentage",
       fill = "Surgery Decision") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c("orange", "royalblue","limegreen"))+
  coord_flip()
```


```{r message=FALSE, warning=FALSE, echo=FALSE}
# Chemo
df_cancer_chemo <- DF %>% 
  dplyr::select(Site,AJCC.7.Stage,Chemotherapy,Study.ID) %>% 
  unite("Cancer_Stage",Site,AJCC.7.Stage,sep = "-") %>% 
  dplyr::group_by(Cancer_Stage,Chemotherapy) %>% 
  dplyr::summarise(num = n()) %>% ungroup() %>% 
  dplyr::arrange(desc(num))
```

```{r echo=FALSE}
df_cancer_chemo %>% 
  mutate(percent = num/sum(num)) %>% 
  mutate(Cancer_Stage = reorder(Cancer_Stage,percent)) %>% 
  top_n(10,wt = percent) %>% 
  ggplot() + 
  geom_bar(aes(x = Cancer_Stage,y = percent, fill = Chemotherapy),stat = "identity") + 
  labs(title = "Top 10 Cancers by Chemo Status", 
       x = "Cancer with Stage",y = "Percentage",
       fill = "Chemo Status") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c("#003399","#00CC33"))+
  coord_flip()
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Radiation
df_cancer_radi <- DF %>% 
  dplyr::select(Site,AJCC.7.Stage,Radiation,Study.ID) %>% 
  unite("Cancer_Stage",Site,AJCC.7.Stage,sep = "-") %>% 
  dplyr::group_by(Cancer_Stage,Radiation) %>% 
  dplyr::summarise(num = n()) %>% ungroup() %>% 
  dplyr::arrange(desc(num))
```

```{r echo=FALSE}
df_cancer_radi %>% 
  mutate(percent = num/sum(num)) %>% 
  mutate(Cancer_Stage = reorder(Cancer_Stage,percent)) %>% 
  top_n(10,wt = percent) %>% 
  ggplot() + 
  geom_bar(aes(x = Cancer_Stage,y = percent, fill = Radiation),stat = "identity") + 
  labs(title = "Top 10 Cancers by Radiation Status", 
       x = "Cancer with Stage",y = "Percentage",
       fill = "Radiation Status") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c("#99CCCC","#6633FF"))+
  theme(legend.position = "bottom", legend.direction = "horizontal")+
  coord_flip()
```

```{r echo=FALSE}
# Data of Oral Cavity
df_oral <- DF %>% dplyr::filter(Site == "Oral Cavity")
```

## Features by Surgery Performed Status

  To begin with, we would like to take a look at the condition of surgery implementation.

**State**

### Environment Attributes by State

  Household Income, Poverty, Unemployment, Language Isolation and Education are all recorded on county level, which can also be regarded as important attributes of the respondents' `SEER.Registry`, so we would like to explore their distribution by states.
  
  For Poverty, the state of Georgia has the highest poverty rate, over 15% in average. For Unemployment, the unemployment rates are quite similar across those 4 states. For Household Income, the state of Georgia has the lowest household income in average. For Language Isolation, the state of Alaska has the lowest language isolation rate, while the state of California has the highest.
  

```{r message=FALSE, echo=FALSE}
g1 <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(X..Persons.Below.Poverty)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Below Poverty Percentage vs State", x = "State", y = "Below Poverty Percentage")

g2 <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(X..Unemployed.ACS.2013.2017)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Unemployed Percentage vs State", x = "State", y = "Unemployed Percentage")

g3 <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(Median.Household.Income)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Median Household Income vs State", x = "State", y = "Median Household Income")

g4 <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Ave = mean(X..Language.isolation.ACS.2013.2017..households.)) %>% 
  ggplot() +
  geom_bar(aes(x = SEER.Registry, y = Ave), stat = "identity") +
  labs(title = "Language Isolation Percentage vs State", x = "State", y = "Language Isolation Percentage")
```

```{r echo=FALSE}
gridExtra::grid.arrange(g1, g2, g3, g4, nrow = 2)
```

```{r message=FALSE, echo=FALSE}
edu_9 <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Avg = mean(X...9th.Grade.Education))
edu_h <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Avg = mean(X...High.School.Education))
edu_b <- df_oral %>% dplyr::group_by(SEER.Registry) %>% dplyr::summarise(Avg = mean(X...Bachelors.Education))
edu_9$Avg1 <- edu_h$Avg
edu_9$Avg2 <- edu_b$Avg

df_int <- stack(edu_9[,2:4])
df_state_edu <- data.frame(rep(c("Alaska", "California", "Connecticut", "Georgia"), 3), df_int)
colnames(df_state_edu) <- c("State", "values", "education")

state <- c(rep("Alaska", 3), rep("California", 3), rep("Connecticut", 3), rep("Georgia", 3))
edu <- rep(c("9th Grade", "High School", "Bachelor"), 4)
values <- c(2.66, 7.61, 28.99, 9.242076, 16.874751, 32.903706, 4.116577, 9.667285, 38.037514, 5.239321, 14.709081, 27.351167)
```

  For Education, the state of California has the highest proportion of respondents are educated below 9th Grade and the highest proportion of respondents are educated below High School. Meanwhile, the state of Connecticut has the highest proportion of respondents are educated below Bachelor.
  
```{r message=FALSE, echo=FALSE}
df_state_edu2 <- data.frame(state, edu, values)
df_state_edu2 %>% 
  ggplot(aes(fill = edu, x = state, y = values/100)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(title = "Education Status per State", x = "State", y = "Percentage", fill = "Education") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))
  
```

  By examining surgery implementation condition by state, we found that California has the largest population of patients researched in the data. While Alaska and Connecticut have the largest proportion of patients who have performed the surgery.
  
```{r message=FALSE, echo=FALSE}
# Registry State vs. Whether the surgery implemented
p1 <- ggplot(df_oral, aes(x = SEER.Registry, fill = Surgery.Performed.) ) +  
    geom_bar( aes(y = ..count../sum(..count..) ) ) +
    labs(title = "Surgery Implementation by State",
         x = "Registered State",
         y = "Percentage") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred") ) 

p2 <- df_oral %>% dplyr::select(SEER.Registry, Study.ID, Surgery.Performed.) %>% 
  dplyr::group_by(SEER.Registry, Surgery.Performed.) %>% 
  dplyr::summarise(num = n()) %>% 
  ungroup() %>% 
  dplyr::mutate(percent = num/sum(num)) %>% 
  
  ggplot()+
  geom_bar(aes(x = SEER.Registry, y = percent, fill = Surgery.Performed.), position = "fill", stat = "identity") + 
  labs(
         x = "Registered State",
         y = "Percentage") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred") )

gridExtra::grid.arrange(p1,p2)
```

**Race & Gender**

  Before looking at the surgery condition by race and gender, we would like to take a look at the distribution of race by gender. The plot above shows that, more than 60% of observations are white people, and more than 40% of observations are white male, which indicates that there would be an imbalance problem in the data.
    
```{r message=FALSE, warning=FALSE, echo=FALSE}
# Distribution of race by gender
df_oral %>% dplyr::select(Race,Sex,Study.ID) %>% 
    dplyr::group_by(Race,Sex) %>% 
    dplyr::summarise(num = n()) %>% 
    ungroup() %>% 
    dplyr::mutate(percent = num/sum(num)) %>% 
    
    ggplot() +
    geom_bar(aes(x = Race,y = percent,fill = Sex),stat = "identity") +
    labs(title = "Distribution of Race by Gender",x = "Race", y = "Percentage") + 
    scale_fill_manual("Gender", values = c("pink", "skyblue"))+
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 10, hjust = 0.5, vjust = 0.5))
```
    
  For surgery condition, the plot below shows that Asian of Pacific Islander has the largest proportion of performing the surgery while the Black has the smallest one. Female and Male has almost the same proportion to accept the surgery, which seems that there is no gender biases.
    
```{r message=FALSE, echo=FALSE}
# Race vs. Whether the surgery implemented

p3 <- df_oral %>% dplyr::select(Race, Study.ID, Surgery.Performed.) %>% 
  dplyr::group_by(Race, Surgery.Performed.) %>% 
  dplyr::summarise(num = n()) %>% 
  ungroup() %>% 
  dplyr::mutate(percent = num/sum(num)) %>% 
  
  ggplot()+
  geom_bar(aes(x = Race, y = percent, fill = Surgery.Performed.), position = "fill", stat = "identity") + 
  labs(title = "Surgery Implementation by Race & Gender",
         x = "Race",
         y = "Percentage") +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme(axis.text.x = element_text(angle = 10, hjust = 0.5, vjust = 0.5)) + 
    guides(fill = F)

# Gender vs. Whether the surgery implemented

p4 <- df_oral %>% dplyr::select(Sex, Study.ID, Surgery.Performed.) %>% 
  dplyr::group_by(Sex, Surgery.Performed.) %>% 
  dplyr::summarise(num = n()) %>% 
  ungroup() %>% 
  dplyr::mutate(percent = num/sum(num)) %>% 
  
  ggplot()+
  geom_bar(aes(x = Sex, y = percent, fill = Surgery.Performed.), position = "fill", stat = "identity") + 
  labs(
         x = "Gender",
         y = "Percentage") +
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))

gridExtra::grid.arrange(p3,p4)
```

  After looking at distribution by gender and race separately, we would like to put these 2 factors together to explore the distribution of performed surgery.
  
  Then, we find that the white male has the largest proportion of performing a surgery. 
    
```{r echo=FALSE}
# Gender cross Race  vs.  Whether the surgery implemented
df2 <- df_oral[,c("Sex","Race", "Surgery.Performed.")] 
df2$Sex <- revalue(df2$Sex, c("Female"="F"))
df2$Sex <- revalue(df2$Sex,c("Male"="M"))
df2$Race <- revalue(df2$Race,c("Asian or Pacific Islander"="Asian/PI"))
df2$Race <- revalue(df2$Race,c("American Indian/Alaska Native"="Indigen"))
df2 <-  unite(df2, "Sex_Race", Sex, Race, sep = "_", remove = FALSE)
```


```{r echo=FALSE}
ggplot(df2, aes(x = Sex_Race, fill = Surgery.Performed.) ) +  
    geom_bar( aes(y = ..count../sum(..count..) ) ) +
    labs(title = "Surgery Implementation by Gender_Race",
         x = "Gender_Race",
         y = "Percentage")+
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) + coord_flip()
```
**Age and Year of Diagnosis**

  The age feature is recorded as a continuous variable, so we make a density plot instead of the bar plot. In the below plot, we find that the distributions of performing surgery and not performing surgery on ages are mostly overlapping. However, we can still see that, for people older than 75 years old, they are more likely to perform the surgery; while for people younger than 75 years old, they are more likely to not perform the surgery.
 
```{r message=FALSE, echo=FALSE}
ggplot(data=df_oral, aes(x=Age.at.Diagnosis, group=Surgery.Performed., fill=Surgery.Performed.)) +
    geom_density(adjust=1.5, alpha=.4) + 
  labs(title = "Surgery Performed Status by Age",
       x = "Age at Diagnosis",y = "Density",
       fill = "Surgery Performed")
```

  Then, we would like to explore the change of surgery implementation on time series. From the plot below, for the trend of amount of performed surgery, we can see that as time goes by, there is an upper trend.

```{r message=FALSE, warning=FALSE, echo=FALSE}
df_oral %>% 
    dplyr::select(Year.of.Diagnosis,Surgery.Performed.,Study.ID) %>% 
    mutate(Year.of.Diagnosis = as.character(Year.of.Diagnosis)) %>% 
    dplyr::group_by(Year.of.Diagnosis,Surgery.Performed.) %>% 
    dplyr::summarise(num = n()) %>% ungroup() %>% 
    ggplot() + 
    geom_line(aes(x = Year.of.Diagnosis,
                  y = num,
                  group = Surgery.Performed.,
                  col = Surgery.Performed.),size = 1) +
    labs(title = "Performed Surgery Trend by Year",
         x = "Year of Dignosis",y = "Count") +
    scale_color_manual("Surgery Implementation",values = c("grey55", "darkred"))
```

**Insurance**

  Besides of the demographic features, we also concern about the impact of social biases like insurance on whether to perform the surgery or not. From below chart, we can see that the large majority of repondents are insured and insured people are most likely to perform the surgery.

```{r message=FALSE, warning=FALSE, echo=FALSE}
p5 <- df_oral %>% dplyr::select(Insurance,Surgery.Performed.,Study.ID) %>% 
    dplyr::group_by(Insurance,Surgery.Performed.) %>% 
    dplyr::summarise(num = n()) %>% 
    ungroup() %>% 
    mutate(percent = num/sum(num)) %>% 
    ggplot() +
    geom_bar(aes(x = Insurance,y = percent,fill = Surgery.Performed.),stat = "identity") +
    labs(title = "Surgery Implementation by Insurance Type",x = NULL, y = "Percentage") + 
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) + coord_flip()

p6 <- df_oral %>% dplyr::select(Insurance,Surgery.Performed.,Study.ID) %>% 
    dplyr::group_by(Insurance,Surgery.Performed.) %>% 
    dplyr::summarise(num = n()) %>% 
    ungroup() %>% 
    mutate(percent = num/sum(num)) %>% 
    ggplot() +
    geom_bar(aes(x = Insurance,y = percent,fill = Surgery.Performed.), position = "fill", stat = "identity") +
    labs(x = NULL, y = "Percentage") + 
    scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
    scale_fill_manual("Surgery Implementation", values = c("grey55", "darkred")) + coord_flip()

gridExtra::grid.arrange(p5,p6)
```


# Feature Engineering

## Feature Establishment

  The columns `Surgery.Performed.`,`Radiation` and `Chemotherapy` contain information of the actual therapy that the patients have taken. It would be more informative to combine these 3 columns together to reflect patients' treatment as a whole. Hence, we create a column `True_Therapy` to get the aggregated information.
  
```{r message=FALSE, warning=FALSE, echo=FALSE}
DF_pre <- df_oral

# Combine variables of Surgery Performed, Chemotherapy and Radiation into one varibale called True_Therapy

DF_pre$rad = ifelse(DF_pre$Radiation =="No radiation and/or cancer-directed surgery", "No", "Yes")
DF_pre$True_Therapy = rep(NA,nrow(DF_pre))

for (i in 1:nrow(DF_pre)){
  
  if(DF_pre$Surgery.Performed.[i] == "Yes" && DF_pre$Chemotherapy[i] == "Yes" && DF_pre$rad[i] == "Yes"){
    DF_pre$True_Therapy[i] = "S+C+R"
  }
  else if(DF_pre$Surgery.Performed.[i] == "Yes" && DF_pre$Chemotherapy[i] == "Yes" && DF_pre$rad[i] == "No"){
    DF_pre$True_Therapy[i] = "S+C"
  }
  else if(DF_pre$Surgery.Performed.[i] == "Yes" && DF_pre$Chemotherapy[i] == "No/Unknown" && DF_pre$rad[i] == "Yes"){
    DF_pre$True_Therapy[i] = "S+R"
  }
  else if(DF_pre$Surgery.Performed.[i] == "Yes" && DF_pre$Chemotherapy[i] == "No/Unknown" && DF_pre$rad[i] == "No"){
    DF_pre$True_Therapy[i] = "S"
  }
  else if(DF_pre$Surgery.Performed.[i] == "No" && DF_pre$Chemotherapy[i] == "No/Unknown" && DF_pre$rad[i] == "No"){
    DF_pre$True_Therapy[i] = "No treatment"
  }
  else if(DF_pre$Surgery.Performed.[i] == "No" && DF_pre$Chemotherapy[i] == "No/Unknown" && DF_pre$rad[i] == "Yes"){
    DF_pre$True_Therapy[i] = "R"
  }
  else if(DF_pre$Surgery.Performed.[i] == "No" && DF_pre$Chemotherapy[i] == "Yes" && DF_pre$rad[i] == "No"){
    DF_pre$True_Therapy[i] = "C"
  }
  else if(DF_pre$Surgery.Performed.[i] == "No" && DF_pre$Chemotherapy[i] == "Yes" && DF_pre$rad[i] == "Yes"){
    DF_pre$True_Therapy[i] = "C+R"
  }
  
}
```

  Meanwhile, the 2021 version **NCCN Guideline** has also provided the information of what kinds of treatments patients with oral cavity cancer are supposed to take. Since the recommended treatment is based on `AJCC.7.Stage`, where contains information of tumor size (`Size`) and lymph nodes status (`Lymph.Nodes`), we will create a column `Rec_Therapy` indicating the treatments that a patient is supposed to have under **NCCN Guideline**.
  
```{r message=FALSE, warning=FALSE, echo=FALSE}
# Create the canser stage for oral cavity
df_recommend <- DF_pre %>% 
  dplyr::select(Site,Size,AJCC.7.Stage,Study.ID,True_Therapy) %>%
  dplyr::filter(Site=="Oral Cavity") %>%
  unite("Cancer_Stage",Site,AJCC.7.Stage,sep = "-") %>% 
  mutate(Cancer_Stage = factor(Cancer_Stage,levels = c("Oral Cavity-I","Oral Cavity-II","Oral Cavity-III","Oral Cavity-IVA","Oral Cavity-IVB","Oral Cavity-IVC","Oral Cavity-IVNOS")))
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Create the variable Rec_Therapy that shows the therapy combinations for patients based on the NCCN guideline.
df_recommend$Rec_Therapy <- rep(NA,nrow(df_recommend))
for(i in 1:nrow(df_recommend)){
  if(df_recommend$Cancer_Stage[i]=="Oral Cavity-I"){
    df_recommend$Rec_Therapy[i]="S+R"
  }
  else if(df_recommend$Cancer_Stage[i]=="Oral Cavity-II"){
    df_recommend$Rec_Therapy[i]="S+R"
  }
  else{
    df_recommend$Rec_Therapy[i]="S"
  }
}
```

  For equity issue, we believe it would be more important to examine the matchness of recommended therapies and the real therapies that patients actually get. So, we create a column `Match` to evaluate such matchness. `Match` is a binary feature, it equals to 1 when `True_Therapy` and `Rec_Therapy` are completely matched or `Rec_Therapy` is the subset of `True_Therapy`(eg. A patient is supposed to take surgery based on NCCN Guideline, and this patient actully has taken surgery and chemo), and it equals to 0 otherwise.
  
```{r message=FALSE, warning=FALSE, echo=FALSE}
# Create the variable Match with binnary outcomes.
df_recommend$Match = ifelse(df_recommend$True_Therapy == df_recommend$Rec_Therapy, 1,0)
for(i in 1:nrow(df_recommend)){
  if(df_recommend$True_Therapy[i]=="S+R" && df_recommend$Rec_Therapy[i] == "S"){
    df_recommend$Match[i] = 1
  }
  else if(df_recommend$True_Therapy[i]=="S+C+R" && df_recommend$Rec_Therapy[i] == "S"){
    df_recommend$Match[i] = 1
  }
  else if(df_recommend$True_Therapy[i]=="S+C+R" && df_recommend$Rec_Therapy[i] == "S+R"){
    df_recommend$Match[i] = 1
  }
}
```


```{r echo=FALSE}
# Merger the data for modeling
oral_cancer <- left_join(DF_pre[,-which((names(DF_pre) %in% c("Site","Subsite","AJCC.7.Stage","Size","Lymph.Nodes","Mets","Surgery.Performed.","Surgery.Decision","Radiation","Chemotherapy","rad","True_Therapy")))],df_recommend,by = "Study.ID")

oral_cancer <- 
  oral_cancer %>% dplyr::rename(
  `X9th.Education` = X...9th.Grade.Education,
  HS.Education = X...High.School.Education,
  Bachelors.Education = X...Bachelors.Education,
  Below.Poverty = X..Persons.Below.Poverty,
  Unemployed = X..Unemployed.ACS.2013.2017,
  Language.Isolation = X..Language.isolation.ACS.2013.2017..households.)

# Data set for modeling
# write.csv(oral_cancer,"oral_cancer_data.csv",row.names = F)
```

  We also do some feature transformations. For the `Median.Household.Income`, we change it from continuous to ordinal with 3 levels. The first level is under first quantile of the income, the second level is between first quanntile and third quantile of the income, and the third level is greater than third quantile of the highest income. For the `Race`, we use the white as the reference group. 
  
  Referring to the below plot, different insured conditions have overlapping distributions on the income population from low to high. So we also change the `Insurance` to be binary with 1 representing the patient who had all kinds of insurance and 0 representing the patient who didn't have any insurance. We also change the `Sex` variable to be binary with 1 representing male and 0 representing female.

```{r echo=FALSE}
oral <-  oral_cancer

break1 <-  c(21840, 57160, 79640, 106760)
label <-  c("Low","Mid","High")
oral$Income_Level <-  oral$Median.Household.Income
oral$Income_Level <- cut(oral$Income_Level, break1, label,include.lowest = T)
```

```{r echo=FALSE}
ggplot(data=oral, aes(x=Median.Household.Income, group=Insurance, fill=Insurance)) +
    geom_density(adjust=1.5, alpha=.4) + 
  labs(title = "Insurance Distribution on Household Income",
       x = "Househole Income",y = "Density",
       fill = "Insurance")
```

```{r echo=FALSE}
oral$Sex <-  ifelse(oral$Sex == "Male", 1, 0)
oral$Race <-  factor(oral$Race, levels = c("White", "Asian or Pacific Islander", "Black", "Hispanic","American Indian/Alaska Native"))
oral$Insurance <- ifelse(oral$Insurance == "Uninsured",0,1)
```

## Feature selection

  When we choose the predictors that will be used in the model, we make a random forest-based selection to see the importance level of all variables. We decide not to use the `Income_Level` since its importance level is relatively low.

```{r echo=FALSE}
# Train-Test Split
set.seed(1)
split_index <-  sort(sample(nrow(oral), nrow(oral)*0.8))
trainset <-  oral[split_index,]
testset <-  oral[-split_index,]
```

```{r echo=FALSE}
# Randomm Forest-based selection
set.seed(1)
rf_select <-  randomForest(as.factor(Match) ~ Bachelors.Education + Sex + Race+ Insurance + Below.Poverty + Unemployed + Language.Isolation + Income_Level + Sex:Race + Cause.of.Death + SEER.Registry + X9th.Education + HS.Education + Age.at.Diagnosis + Cancer_Stage, data = trainset, mtry = 4, importance = T)

# Visualize the importance
# importance(rf_select)
varImpPlot(rf_select)
```

# Model

## Logistic Regression

  Since the outcome is binary, firstly we would like to use the logistic regression model. Except for all the predictors that we decide to use, we also add the interaction term of `Race` and `Sex`.

```{r echo=FALSE}
m1 <-  glm(Match ~ Cancer_Stage  + Language.Isolation + X9th.Education + Bachelors.Education + Unemployed + Age.at.Diagnosis + Below.Poverty + Race + Sex + Race:Sex + Insurance, data = trainset, family = binomial(link = "logit"))

summary(m1)

coe = coef(m1)
invlogit(coe)
```

**Interpretation**

  From the summary of the logistic regression, we can see that all of the estimated coefficients for `Cancer_Stage` are significant. Especially for stage III, patients are more likely to follow the guideline of therapy compared with other stages. The estimated coefficient for `Age.at.Diagnosis`, `RaceBlack`, `Sex` and `Insurance` are all significant, and the first three coefficients are negative which means that the older patients, blacks and males tend to not follow the guideline of therapy. The estimated coefficient for `Insurance` is positive and significant which means that patients who have insurance are more likely to follow the guideline of therapy than those who don't have insurance. From the prospective of porbability, the blacks are 37% less likely to get correct therapy than that of whites. The patients with no insurance are 63% less likely to follow the guideline of therapy than patients who have insurance. Females are 18% more likely to follow the guideline of therapy than males.
  
**Binned Residual plot**

  From the binned residual plot, we can see that lots of points are outside the bin which needs further adjustment.

```{r echo=FALSE}
# Residual Plor-Logistic Model
binnedplot(fitted(m1), residuals(m1))
```

**Confusion Matrix**

  According to the confusion matrix about the test dataset, we see that the prediction accuracy rate is about 80.17% which is pretty high.

```{r echo=FALSE}
logis_test_pred <- factor(ifelse(predict(m1, newdata = testset,type = "response") < 0.5, 0, 1))

table(logis_test_pred,testset$Match)

logis_acc_base <- (690+430)/(690+141+136+430)
logis_acc_base
```

**Cross-Validation**

  To test the sensitivity of logistic model, we use the K-fold cross validation to evaluate the change of accuracy as we use different train set and test set. The result shows that, the accuracy will change dramatically if we adjust the dataset to test the logistic model.
  
```{r echo=FALSE}
# K-Fold Validation
set.seed(7)
folds <- createFolds(y=oral$Match,k=10)
logis_cv <- data.frame(number = rep(NA,10), test_accuracy = rep(NA,10), train_accuracy = rep(NA,10))
```

```{r echo=FALSE}
# 10-fold CV
for(i in 1:10){
  
 fold_test <- oral[folds[[i]],]
 fold_train <- oral[-folds[[i]],]
  
  
 fold_model <- glm(Match ~ Cancer_Stage  + Language.Isolation + X9th.Education + Bachelors.Education + Unemployed + Age.at.Diagnosis + Below.Poverty + Race + Sex + Race:Sex + Insurance, data = fold_train, family = binomial(link = "logit"))
 
 # Test Accuracy
 fold_predict <- predict(fold_model,type='response',newdata=fold_test)
 fold_predict <- ifelse(fold_predict > 0.5,1,0)
 
 fold_test$predict <-  fold_predict
 fold_test_error <-  mean(fold_test$predict!=fold_test$Match)
 fold_test_accuracy <-  1 - fold_test_error
 logis_cv$number[i] <-  paste0("k=",i)
 logis_cv$test_accuracy[i] <-  fold_test_accuracy
 
 # Train Accuracy
 fold_predict2 <- predict(fold_model,type="response",newdata=fold_train)
 fold_predict2 <- ifelse(fold_predict2 > 0.5,1,0)
 fold_train$predict <-  fold_predict2
 fold_train_error <-  mean(fold_train$predict!=fold_train$Match)
 fold_train_accuracy <-  1- fold_train_error
 logis_cv$train_accuracy[i] <-  fold_train_accuracy
}
```

```{r echo=FALSE}
# Visualize K-fold CV
logis_cv %>% mutate(number = factor(number,levels = c("k=1","k=2","k=3","k=4","k=5","k=6","k=7","k=8","k=9","k=10"))) %>% 
  ggplot() + 
  geom_line(aes(x = number,y = test_accuracy,group = 1,col = "Test"))+
  geom_line(aes(x = number,y = train_accuracy,group = 1,col = "Train")) + 
  labs(title = "K-fold Cross Validation: Logistic Model",
       subtitle = "K = 10",
       x = NULL,y = "Accuracy",col = "Accuracy")
```

## Naive Bayes

  Naive Bayesian classifier can take the given variables as conditions to calculate the probability that an observation can be classified into a certain category. 
  
  In this project, we have a lot of features that relates to the matching between the therapy a patient is supposed to have and actually have. These features can be seen as given conditions, which are also known as prior information. So, Naive Bayesian classifier would be appropriate for this problem.
  
  By using `e1071` package, we have a initial Naive Bayesian model with an accuracy of 79.81%, which is slightly lower than the accuracy of the logistic (80.17%).
  
  When looking at the conditional probability in the result of the model, we can see that the probability of matching given $Race=Black$ equals to 8.1%, and probability of matching given $Race=White$ equals to 66.7%, so there is large discrepancy between these two groups. However, white people also have the highest probability of mismatching, which might due to the imbalanced white population in the dataset.
  
```{r echo=FALSE}
m2 <- naiveBayes(as.factor(Match) ~ Cancer_Stage  + Language.Isolation + X9th.Education + Bachelors.Education + Unemployed + Age.at.Diagnosis + Below.Poverty + Race + as.factor(Sex) + as.factor(Insurance),data = trainset)

bayes_test_pred <- predict(m2, newdata = testset) 
```

```{r echo=FALSE}
kableExtra::kable(m2$tables$Race)
```

```{r echo=FALSE}
table(bayes_test_pred,as.factor(testset$Match))
1-mean(bayes_test_pred!=as.factor(testset$Match))
```

  Then we do the tuning of the `laplace` parameter, the red line indicates the accuracy of the initial model, and the blue line indicates the accuracy of the logistic model. When $laplace=100$, the model reached the highest accuracy which might due to the fact that this model takes conditional probabilities into consideration.
  
```{r echo=FALSE}
laplace_value <- seq(0,200,by=10)
train_err_bye <- rep(NA, length(laplace_value))

set.seed(1)
for (i in 1:length(laplace_value)) {
    bye_fold_model <-naiveBayes(as.factor(Match) ~ Cancer_Stage  + Language.Isolation + X9th.Education + Bachelors.Education + Unemployed + Age.at.Diagnosis + Below.Poverty + Race + as.factor(Sex )+ as.factor(Insurance),data = trainset,laplace = laplace_value[i])
    
    bye_fold_pred <- predict(bye_fold_model,testset,laplace = laplace_value[i])
    train_err_bye[i] <- mean(bye_fold_pred!=as.factor(testset$Match))
}
```

```{r echo=FALSE}
ggplot(data = data.frame(laplace = laplace_value, err = 1-train_err_bye))+
  geom_point(aes(x = laplace, y = err)) + 
  geom_line(aes(x = laplace, y = err)) + 
  geom_hline(yintercept = 1-mean(bayes_test_pred!=as.factor(testset$Match)),linetype = 2, col="red") +
   geom_hline(yintercept = logis_acc_base,linetype = 2, col="darkblue")+
  scale_x_continuous(breaks = seq(0,200,by=10)) +
  labs(title = "Tuning: Laplace parapmeter in Naive Bayes on Test Set",x = "Laplace",y = "Accuracy") + 
  theme_bw()
```

# Discussion

  Firstly, the dataset itself has contained bias that the majority of the respondents are insured white male from California which result in imbalance. Secondly, the binned residual plot suggests that we still need further research to make the plot look better. Thirdly, about the Naive Bayes, we should have also use k fold cross-validation on the Naive Bayes model, but its computation was extremely time-consuming. Besides, the hypothesis of Naive Bayesian classifier is that all variables' impact on the outcome is independent, but there are interactions among our features. Last but most importantly, there are lots of uncertainties in the dataset which are usually represented by "unknown". For example, in the `Radiation` columns, for many respondents, whether the radiation conducted before or after the surgery is unknown. Regardless of these uncertainties, we can research whether the order between radiation and surgery has impact on the therapy or not.


# Conclusion

  In general, we can conclude that there exists bias in this data since we know that blacks are less likely to get correct therapy compared with other races from the result of the model. We can also see that patients with no insurance are less likely to take treatments. Besides, there are some subtle trends that patients with low educational level and patients who are unemployed are slightly tend to not follow the guideline of therapy.

# Reference

(1) Hadley Wickham (2011). The Split-Apply-Combine Strategy for Data Analysis. Journal of Statistical Software, 40(1), 1-29. URL http://www.jstatsoft.org/v40/i01/.

(2) Hadley Wickham (2007). Reshaping Data with the reshape Package. Journal of Statistical Software, 21(12), 1-20. URL http://www.jstatsoft.org/v21/i12/.

(3) Hadley Wickham (2020). tidyr: Tidy Messy Data. R package version 1.1.2. https://CRAN.R-project.org/package=tidyr

(4) Simon Garnier (2018). viridis: Default Color Maps from 'matplotlib'. R package version 0.5.1. https://CRAN.R-project.org/package=viridis

(5) Revelle, W. (2020) psych: Procedures for Personality and Psychological Research, Northwestern University, Evanston, Illinois, USA, https://CRAN.R-project.org/package=psych Version = 2.1.3,.

(6) Goodrich B, Gabry J, Ali I & Brilleman S. (2020). rstanarm: Bayesian applied regression modeling via Stan. R package version 2.21.1 https://mc-stan.org/rstanarm.

(7) Brilleman SL, Crowther MJ, Moreno-Betancur M, Buros Novik J & Wolfe R. Joint longitudinal and time-to-event models via Stan. StanCon 2018. 10-12 Jan 2018. Pacific Grove, CA, USA. https://github.com/stan-dev/stancon_talks/

(8) Andrew Gelman and Yu-Sung Su (2020). arm: Data Analysis Using Regression and Multilevel/Hierarchical Models. R package version 1.11-2. https://CRAN.R-project.org/package=arm

(9) A. Liaw and M. Wiener (2002). Classification and Regression by randomForest. R News 2(3), 18--22.

(10) Max Kuhn (2020). caret: Classification and Regression Training. R package version 6.0-86. https://CRAN.R-project.org/package=caret

(11) David Meyer, Evgenia Dimitriadou, Kurt Hornik, Andreas Weingessel and Friedrich Leisch (2020). e1071: Misc Functions of the Department of Statistics, Probability Theory Group (Formerly: E1071), TU Wien. R package version 1.7-4. https://CRAN.R-project.org/package=e1071

(12) National Comprehensive Cancer Network. (2021). Available from https://www.nccn.org/
